% sections/abstract.tex

\begin{abstract}
Large language models are increasingly used as automated judges to evaluate generated text, yet they are overconfident, positionally biased, and sensitive to evaluation-protocol wording.
We study \emph{protocol (in)stability} as a per-instance reliability signal for LLM judges---whether a verdict survives semantically equivalent rubric paraphrases, response-order permutations, and stochastic resampling---and establish its central and previously uncharacterized property: \textbf{the signal is competence-gated and inverts below a judge-capability threshold.}
Above the threshold, instability cleanly predicts error (accuracy gaps up to 27 points between stable and unstable subsets on a mid-capability judge); below it, a near-random judge produces \emph{stable-but-wrong} verdicts (65\% of Qwen2.5-1.5B's JudgeBench errors are rubric-stable), so stability must not be read as correctness.
A negative-control experiment confirms the mechanism: on a mid-capability judge, intentionally non-equivalent rubrics flip verdicts $2.2\times$ more often than equivalent paraphrases, whereas a near-random judge is insensitive to the distinction.
We operationalize this finding as \textbf{CARE-Judge} (Calibrated, Abstaining, and Robust Evaluation), which fuses protocol-stability features into a learned correctness calibrator paired with an inherited empirical selective-risk procedure (a held-out threshold targeted by an exact Clopper--Pearson bound, validated by measured violation rates rather than a new theorem).
We evaluate across five judges (a Qwen2.5-1.5B/7B ladder plus a controlled 1.5B/7B/14B size ablation, DeepSeek-V4, GPT-5.5) and four benchmarks (JudgeBench, TL;DR, RewardBench, LMArena) totaling over 22{,}000 pairwise evaluations. Fusion gives its largest gains in the mid-capability regime the mechanism predicts (up to $+9.8$ AUROC over the best single signal on DeepSeek-V4, replicated on Qwen2.5-7B), ties already-strong base confidence for a frontier judge, and adds nothing where a judge is near-chance on a task. The mid-capability effect, tested on a regime fixed a priori by judge accuracy, is the primary result and is positive on all eight pairs; the pooled 17-pair average is diluted by regimes where fusion cannot help by construction.
In selective terms it abstains entirely where a judge is near-random and, where the judge is competent but imperfect, trades coverage for accuracy under an explicit risk budget.
An anonymized code and feature-trace archive accompanies the submission.
\end{abstract}