% sections/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

We studied \emph{protocol-stability} as a per-instance uncertainty signal for LLM-as-a-judge: a verdict that changes under semantically equivalent rubrics, response orderings, or stochastic resampling is often unreliable. CARE-Judge fuses rubric-paraphrase instability, position-swap inconsistency, and self-consistency into a learned correctness calibrator and pairs it with established selective-risk control. We are explicit that the statistical machinery is adopted from prior selective-judging work~\citep{jung2025trust,badshah2026scope}; our contribution is the multi-view protocol-stability representation and a characterization of when it works.

Our findings answer the three research questions. (RQ1) Protocol instability separates correct from incorrect verdicts by 12--54pp, but only above a minimum judge competence; the direction reverses on a near-random judge. (RQ2) At a matched six-call budget, fusion attains the best held-out AUROC in 7 of 9 judge--dataset pairs, beating base confidence, self-consistency, rubric stability, and the SCOPE-style bidirectional signal. (RQ3) The framework's most informative failure is the \emph{stable-but-wrong} regime (Qwen-1.5B on JudgeBench), where consistent verdicts arise from a shared erroneous heuristic; there, safe abstention is the correct behavior but should not be mistaken for a performance win.

\paragraph{Limitations.}
Our guarantees are of the standard split-conformal kind and assume exchangeability between calibration and test items; they can degrade under distribution shift, and they require reference labels to fit the calibrator and select the threshold, which is a real annotation cost. The core claim is asymmetric: stability does not imply correctness, and a judge with a shared bias can defeat every signal simultaneously. Our datasets are modest in size (TL;DR/MT-Bench: 300; JudgeBench: 620), so calibration and test splits are small and low-coverage cells have high variance; we report per-seed risk and violation rates rather than only means. The three judges differ along several axes at once, so cross-model trends are capability-suggestive, not controlled size effects. The cascade and cross-dataset transfer results (Appendices~\ref{app:cascade}--\ref{app:transfer}) now use the same strict three-way split as the main tables, with conditional per-tier calibration and a union-bound error budget, but we still label them exploratory because they rely on a single labeled pool per dataset and have not been stress-tested under distribution shift.

\paragraph{Future Work.}
The most important next steps are a fully budget-matched comparison against Trust-or-Escalate and SCOPE, a conditionally-calibrated cascade with a union-bound error budget and an accompanying guarantee, calibration-size sensitivity curves, and larger-scale evaluation with a controlled open-model capability ladder to isolate capability from size. Extending protocol-stability signals to pointwise (non-pairwise) scoring and to human-annotation triage are natural applications.
