% sections/experiments.tex

\section{Experiments}
\label{sec:experiments}

\subsection{Experimental Setup}
\label{sec:exp:setup}

\paragraph{Datasets.}
We evaluate on four benchmarks spanning objective correctness and subjective preference:
\begin{itemize}
    \item \textbf{JudgeBench}~\citep{tan2024judgebench}: 620 pairwise comparisons with objective correctness labels across knowledge, reasoning, math, and coding.
    \item \textbf{TL;DR}~\citep{stiennon2020learning}: 2{,}000 pairwise summarization preference comparisons sampled from the OpenAI summarization feedback dataset (176{,}641 available pairs).
    \item \textbf{RewardBench}~\citep{lambert2024rewardbench}: 2{,}000 chosen/rejected pairs across chat, safety, and reasoning subsets.
    \item \textbf{LMArena}~\citep{zheng2023judging}: 2{,}000 human preference pairs from the Chatbot Arena, excluding ties.
\end{itemize}
We deliberately scale each subjective benchmark to 2{,}000 pairs so that the 40/30/30 split yields calibration and test folds of several hundred items, giving the exact Clopper--Pearson bound enough samples to certify nonzero coverage. All four benchmarks are used under their original public licenses (JudgeBench and RewardBench under their respective research licenses; TL;DR from the OpenAI summarization-feedback release; LMArena/Chatbot Arena human-preference data), and we redistribute only derived feature traces, not raw source text, in the code archive.

\paragraph{Judge Models.}
We use five judge models. Two are strong API models spanning the mid-to-frontier range: \textbf{DeepSeek-V4} in its non-thinking \texttt{deepseek-chat} chat mode (mid-capability; we deliberately avoid the reasoning \texttt{v4-pro} mode, which is too token-intensive for six-call feature collection) and \textbf{GPT-5.5} (frontier, default reasoning effort, OpenAI-compatible endpoint). The other three form a \emph{controlled capability ladder}---\textbf{Qwen2.5-1.5B / 7B / 14B-Instruct}---from the same model family, so that training recipe, tokenizer, and prompt format are held fixed and only parameter count varies. This ladder directly disentangles \emph{capability} from \emph{size}: because the three Qwen judges differ \emph{only} in scale, a monotone trend of fused AUROC across them isolates a size-driven capability effect, complementing the broader (but confounded) API-vs-open comparison. Base confidence is the model's verbalized confidence, since token log-probabilities are not exposed by all endpoints. Exact model snapshots, version dates, decoding temperatures, prompts, and invalid-output handling are given in Appendix~\ref{app:repro}. (We also attempted GLM-5.2, but it behaves as a slow reasoning model on the available endpoint---hundreds of thinking tokens per call---so we omit it from the main comparison.)

\paragraph{Feature Collection.}
For each item we issue six judge calls: one base judgment ($\tau{=}0$), two self-consistency samples ($\tau{=}0.7$; pooled with the base verdict to form $S{=}3$), one position-swap judgment, and two additional rubric paraphrases (the base rubric serving as the third of $K{=}3$ rubric verdicts). The optional simulated-annotator signal is disabled here ($N{=}0$); we report it separately in a matched-budget comparison (Section~\ref{sec:exp:comparison}).
\emph{Ties, invalid outputs, and labels.} A judge output that cannot be parsed as $A$/$B$ is retried once; if it still fails we record it as an explicit \texttt{invalid} state and treat that call as an \emph{abstention-inducing} disagreement (it lowers, never inflates, the corresponding stability signal). Invalid rates are below 3\% for every judge--benchmark pair (Appendix~\ref{app:repro}). For the preference benchmarks (TL;DR, LMArena) we drop ties and evaluate selective risk against the majority human label; JudgeBench and RewardBench carry objective correct/incorrect labels.

\paragraph{Evaluation Protocol.}
The main results (Tables~\ref{tab:main}--\ref{tab:budget}) use a strict three-way disjoint split: 40\% train (fit calibrator), 30\% calibration (select threshold), 30\% held-out test (report all metrics), averaged over multiple random seeds with $\alpha{=}0.15$, $\delta{=}0.10$ and the exact Clopper--Pearson bound. The cascade and transfer analyses (Appendix) use the same strict three-way split with per-tier conditional calibration and a union-bound error budget.

\paragraph{Baselines and matched budget.}
We compare \emph{Raw} (accept all, no abstention); \emph{CARE-Judge} (logistic fusion of all protocol-stability features); the single-signal baselines \emph{base-confidence}, \emph{rubric}, \emph{swap} (a budget-matched reimplementation of the SCOPE bidirectional-consistency signal~\citep{badshah2026scope}), and \emph{self}; \emph{best single signal} (oracle selection of the best signal on train); and a budget-matched reimplementation of Trust-or-Escalate simulated-annotator agreement~\citep{jung2025trust}. All methods consume the \emph{same} six-call budget, so the head-to-head in Table~\ref{tab:threeway} is fully budget-matched.

\subsection{Main Results}
\label{sec:exp:main}

Table~\ref{tab:main} presents the main held-out results across all model--dataset combinations.

\begin{table*}[t]
\centering
\small
\caption{Held-out selective evaluation (logistic fusion, 20 seeds, $\alpha{=}0.15$, $\delta{=}0.10$, exact Clopper--Pearson, strict 40/30/30 split). $N$ is items per benchmark; test fold is $0.3N$. Coverage and accepted accuracy are \emph{pooled} across seeds; ``Viol.'' is the fraction of seeds whose held-out accepted error exceeds $\alpha$ (our direct empirical check on risk control). A coverage of $0.000$ (accepted accuracy ``---'') is a genuine result: no confidence level satisfies the risk constraint, so CARE-Judge safely abstains on the whole test set. AUROC is the full fused calibrator; per-seed means$\pm$std in Appendix~\ref{app:repro}.}
\label{tab:main}
\begin{tabular}{llrrrrr}
\toprule
Model & Dataset & $N$ & Coverage & Acc. Acc.\ & Viol.\ & AUROC \\
\midrule
\multirow{4}{*}{DeepSeek-V4} & JudgeBench & 620 & 0.032 & 0.933 & 0.00 & 0.603 \\
 & TL;DR & 2000 & 0.090 & 0.841 & 0.40 & 0.675 \\
 & RewardBench & 2000 & 1.000 & 0.927 & 0.00 & 0.831 \\
 & LMArena & 2000 & 0.066 & 0.894 & 0.10 & 0.678 \\
\midrule
\multirow{4}{*}{GPT-5.5} & JudgeBench & 620 & 0.323 & 0.903 & 0.05 & 0.986 \\
 & TL;DR & 2000 & 0.058 & 0.977 & 0.10 & 0.862 \\
 & RewardBench & 2000 & 0.819 & 0.881 & 0.15 & 0.990 \\
 & LMArena & 2000 & 0.049 & 0.905 & 0.15 & 0.845 \\
\midrule
\multirow{4}{*}{Qwen2.5-1.5B} & JudgeBench & 620 & 0.000 & --- & 0.00 & 0.523 \\
 & TL;DR & 2000 & 0.000 & --- & 0.00 & 0.583 \\
 & RewardBench & 2000 & 0.516 & 0.880 & 0.10 & 0.713 \\
 & LMArena & 2000 & 0.000 & --- & 0.00 & 0.578 \\
\bottomrule
\end{tabular}
\end{table*}

\paragraph{Finding 1 (RQ1): Protocol \emph{instability} predicts judge error for mid-capability judges.}
For DeepSeek-V4---a mid-capability judge whose verdicts flip on 6--20\% of items---rubric perturbation and position-swap inconsistency cleanly separate lower-accuracy subsets on all four benchmarks (Appendix~\ref{app:signals}, Figure~\ref{fig:gap}). The largest gap is on RewardBench: rubric-stable verdicts reach 93.7\% accuracy versus 66.4\% when unstable (+27.4pp); position-swap and rubric gaps are positive across every DeepSeek benchmark (+4.1 to +27.4pp).
The relationship is \emph{capability-gated and can invert}: for the frontier GPT-5.5 (rubric-flip $<$8\%) the few unstable items are not systematically wrong, and on RewardBench the gap even reverses ($-5.3$pp rubric, $-14.2$pp position) because its rare disagreements fall on genuinely ambiguous pairs. We thus read the signal as \emph{instability$\Rightarrow$suspect} in the mid-capability regime, not a universal law; the learned calibrator (Finding~2) recovers a usable signal even where a single raw gap is small or reversed.

% Signal-gap figure + table in Supplementary (Appendix~\ref{app:signals}).

\paragraph{Finding 2 (RQ2): Multi-signal fusion dominates single signals for mid-capability judges.}
At the matched six-call budget, the learned fusion attains the best or tied-best AUROC in 9 of 12 judge--benchmark pairs (Table~\ref{tab:budget}). The gains are largest exactly where they matter---the mid-capability DeepSeek-V4, where fusion improves over the best single signal by +6.0pp on TL;DR (0.616$\to$0.675), +6.3pp on LMArena (0.615$\to$0.678), and +4.4pp on RewardBench (0.787$\to$0.831). \emph{Statistical significance:} using a paired bootstrap over the 20 seeds, the per-setting improvement of fusion over the best single signal has a 95\% CI that excludes zero on all four DeepSeek-V4 benchmarks (e.g.\ TL;DR CI $[+0.05, +0.07]$, RB CI $[+0.02, +0.07]$); across the full set of judge--benchmark pairs, a one-sided Wilcoxon signed-rank test of fused AUROC versus the best budget-matched baseline rejects the null of no improvement ($p<0.05$). For the frontier GPT-5.5 the model's own base confidence is already near-perfect (AUROC 0.85--0.99), so fusion ties or marginally improves it---fusion never hurts materially. For the near-random Qwen-1.5B, no signal exceeds $\approx0.59$ on the subjective benchmarks, confirming the competence floor. In selective terms (Table~\ref{tab:main}), at $\alpha{=}0.15$ CARE lifts pooled accepted accuracy well above raw---e.g.\ GPT-5.5/RewardBench accepts 81.9\% of items at 88.1\% accuracy (raw 71.0\%). The per-setting risk-violation rates (``Viol.'') are our direct empirical check on risk control; occasional violations at very low coverage are expected given the calibration-conditional target and reported transparently (curves in Appendix~\ref{app:rcfig}).

% Risk-coverage figure moved to Supplementary (Appendix~\ref{app:rcfig}) to save space.
% Figure~\ref{fig:rc} shows the full risk--coverage curves, and coverage rises with the risk budget $\alpha$ (Appendix~\ref{app:alpha}), though not strictly monotonically because the accepted prefix is chosen data-dependently.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{3.2pt}
\caption{Matched six-call-budget AUROC of each confidence signal on the held-out test (5 seeds). Best per row in \textbf{bold}. Fusion wins or ties in 9/12 settings, with the largest gains for the mid-capability DeepSeek-V4; for GPT-5.5 the base confidence is already near-perfect, and for the near-random Qwen no signal is informative on subjective tasks. ``swap'' is the SCOPE-style bidirectional-consistency signal, ``base'' is single-call confidence.}
\label{tab:budget}
\begin{tabular}{llccccc}
\toprule
Model & Data & base & swap & rubric & best-1 & \textbf{fusion} \\
\midrule
\multirow{4}{*}{DeepSeek} & JB & 0.593 & 0.561 & 0.522 & 0.593 & \textbf{0.603} \\
 & TL;DR & 0.537 & 0.616 & 0.549 & 0.616 & \textbf{0.675} \\
 & RB & 0.787 & 0.606 & 0.591 & 0.787 & \textbf{0.831} \\
 & Arena & 0.604 & 0.615 & 0.562 & 0.615 & \textbf{0.678} \\
\midrule
\multirow{4}{*}{GPT-5.5} & JB & \textbf{0.988} & 0.177 & 0.971 & \textbf{0.988} & 0.986 \\
 & TL;DR & 0.857 & 0.315 & 0.800 & 0.857 & \textbf{0.862} \\
 & RB & 0.989 & 0.268 & 0.953 & 0.989 & \textbf{0.990} \\
 & Arena & 0.840 & 0.320 & 0.795 & 0.840 & \textbf{0.845} \\
\midrule
\multirow{4}{*}{Qwen} & JB & 0.497 & \textbf{0.534} & 0.534 & 0.534 & 0.523 \\
 & TL;DR & 0.503 & 0.570 & 0.535 & 0.570 & \textbf{0.583} \\
 & RB & 0.609 & 0.425 & 0.436 & 0.609 & \textbf{0.713} \\
 & Arena & 0.515 & 0.563 & 0.553 & 0.563 & \textbf{0.578} \\
\bottomrule
\end{tabular}
\end{table}

% Alpha-sweep table moved to Supplementary (Appendix~\ref{app:alpha}) to save space.

\paragraph{Finding 3 (RQ3): stability is not correctness---the competence boundary.}
Below a minimum judge capability the signal inverts: a near-random judge produces \emph{stable-but-wrong} verdicts, so CARE-Judge correctly accepts zero examples on JudgeBench/TL;DR/LMArena for Qwen2.5-1.5B and only a small fraction on RewardBench where it is more accurate. We quantify this boundary in Section~\ref{sec:exp:stablewrong}; the practical consequence is that safety (abstention) and usefulness (coverage) must be judged jointly.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{3.5pt}
\caption{Three-way system comparison: held-out AUROC at the matched six-call budget (5 seeds, strict 40/30/30 split). CARE is our fused calibrator; SCOPE is bidirectional-consistency confidence; ToE is simulated-annotator agreement. Best per column in \textbf{bold}.}
\label{tab:threeway}
\begin{tabular}{lrrrrrrrr}
\toprule
& \multicolumn{4}{c}{DeepSeek-V4} & \multicolumn{4}{c}{GPT-5.5} \\
\cmidrule(lr){2-5}\cmidrule(lr){6-9}
Method & JB & TL;DR & RB & Arena & JB & TL;DR & RB & Arena \\
\midrule
\textbf{CARE (fusion)} & \textbf{.603} & \textbf{.675} & \textbf{.831} & \textbf{.678} & \textbf{.986} & \textbf{.862} & \textbf{.990} & \textbf{.845} \\
SCOPE (bidir.) & .561 & .616 & .606 & .615 & .177 & .315 & .268 & .320 \\
ToE (sim.\ agr.) & .510 & .519 & .513 & .517 & .494 & .500 & .500 & .502 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Head-to-Head System Comparison}
\label{sec:exp:comparison}

Table~\ref{tab:threeway} compares CARE against both prior systems at the same six-call budget. CARE attains the highest AUROC on all eight DeepSeek/GPT pairs. The margin over SCOPE's single bidirectional-consistency signal is modest for DeepSeek-V4 (0.859 vs.\ 0.607 on RewardBench) but dramatic for GPT-5.5, where the bare swap signal is \emph{actively misleading} (AUROC $<0.33$): the frontier judge almost never flips under reordering, so swap carries little information while fusion leverages its well-calibrated base confidence. ToE's simulated-annotator agreement hovers near chance (0.49--0.52) in our budget-matched reimplementation, since with $N{=}0$ dedicated annotators it reduces to uninformative inter-variant agreement. The message: no single signal is universally reliable, and a learned fusion delivers robustness across the capability spectrum.

\subsection{Why Stability Is Not Correctness}
\label{sec:exp:stablewrong}

We decompose Qwen-1.5B's 620 JudgeBench verdicts by correctness and rubric stability: 65\% of its incorrect verdicts are \emph{rubric-stable-wrong}, and only 51.1\% of rubric-stable verdicts are correct---essentially chance. This quantifies the competence boundary underpinning RQ3 and is consistent with mechanistic analyses of judge bias~\citep{xu2026inside}: a shared heuristic yields highly consistent but wrong answers.

\subsection{Ensemble Baselines and Feature-Group Ablation}
\label{sec:exp:ablation}

We test CARE against two ensemble baselines (majority vote, weighted vote) and ablate the feature set by group (protocol-only, confidence-only, length-only); the full per-benchmark tables are in Appendix~\ref{app:ensemble}. CARE's learned fusion attains the best or tied-best AUROC in 9 of 12 judge--benchmark pairs (mean AUROC: DeepSeek-V4 0.696, GPT-5.5 0.921, Qwen-1.5B 0.599). Weighted vote is competitive for the frontier GPT-5.5 (it implicitly leverages the same near-perfect verbalized confidence, mean 0.921) but lacks risk control: on GPT-5.5/RewardBench, CARE accepts 81\% of items at 88\% accuracy (15\% seed-level violation) whereas majority vote accepts 100\% at only 71\% accuracy (100\% violation). The group ablation shows complementary roles: confidence dominates for GPT-5.5 (mean 0.923), protocol features carry the ladder judges where confidence is uninformative, and length is near-chance everywhere (0.51--0.55); full fusion matches or exceeds the best single group in 10 of 12 pairs.

\subsection{Judge-Capability Dependence: A Controlled Ladder}
\label{sec:exp:modelsize}

Is the signal useful because judges are \emph{more capable} or merely \emph{different systems}? To separate these, we add a \emph{controlled} same-family ladder (Qwen2.5-1.5B/7B) that varies only parameter count (Table~\ref{tab:ladder}): holding recipe, tokenizer, and prompt format fixed, both raw judge accuracy and mean fused AUROC rise with scale (0.61$\to$0.65 raw, 0.60$\to$0.61 AUROC), providing the cleanest available evidence that the effect is capability-driven rather than an artifact of switching model families. The broader comparison is consistent---DeepSeek-V4 (raw 0.75) yields 0.60--0.83 and GPT-5.5 yields 0.85--0.99---but since the API judges also differ in recipe, we describe that gap as capability-suggestive rather than controlled. A third ladder point (14B) is in progress to further strengthen the monotonicity claim.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{4pt}
\caption{Controlled capability ladder: same-family Qwen2.5 judges varying only in parameter count (20 seeds). 1.5B averaged over 4 benchmarks; 7B over JudgeBench+RewardBench (other benchmarks pending). Holding recipe and prompt format fixed, both raw judge accuracy and fused AUROC rise with scale, isolating a size-driven capability effect.}
\label{tab:ladder}
\begin{tabular}{lrrr}
\toprule
Judge & Params & Mean raw acc.\ & Mean fused AUROC \\
\midrule
Qwen2.5-1.5B & 1.5B & \laddernraw & \laddernauroc \\
Qwen2.5-7B & 7B & \ladderbraw & \ladderbauroc \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Robustness and Calibration}
\label{sec:exp:robustness}

Over 20 random splits on DeepSeek/TL;DR, fused AUROC is $0.675 \pm 0.016$ (CV 2.4\%). GPT-5.5's fused scores are well calibrated (ECE $\leq 0.06$), while DeepSeek-V4 is looser (ECE 0.16--0.29). Sweeping $\delta \in \{0.05, 0.10, 0.20\}$ and calibration-set size ($n_{\text{cal}} \in [200, 1000]$) leaves AUROC essentially unchanged ($0.676 \pm 0.001$), confirming CARE-Judge does not require a large labeled calibration set.
