% sections/introduction.tex

\section{Introduction}
\label{sec:intro}

Large language models (LLMs) are now widely deployed as automated judges---scoring RLHF outputs~\citep{ouyang2022training}, ranking chatbot responses~\citep{zheng2023judging}, and evaluating generation quality---but they make systematic errors, exhibit positional biases, and vary verdicts based on evaluation-protocol wording, all while expressing high confidence~\citep{li2024llmsasjudges,wang2023faireval,wei2024systematic}.

Prior work has established that LLM judges are overconfident and has begun to control this risk explicitly.
\citet{jung2025trust} showed that standard confidence measures (predictive probability, verbalized confidence) poorly predict whether a judge's verdict agrees with human annotators, and proposed \emph{Simulated Annotators} with cascaded selective evaluation to provide provable human-agreement guarantees; \citet{badshah2026scope} instead aggregate both response orderings into a permutation-invariant conformal score.
These methods each rely on a \emph{single} confidence source---simulated-annotator agreement or bidirectional preference probability---and do not exploit the broader observation that sensitivity across \emph{multiple} semantically equivalent evaluation protocols is itself an error signal.
We build directly on their selective-risk machinery (a held-out calibration threshold with an exact Clopper--Pearson bound) and do not claim it as a contribution; our contribution is the confidence signal that machinery is applied to. We are careful \emph{not} to over-claim its formal guarantees: because our acceptance threshold is chosen data-dependently on the calibration split, we treat the risk criterion as an empirical target validated by measured violation rates, not as a family-wise-corrected sequential test.

We investigate a complementary and deliberately asymmetric hypothesis: \textbf{a judgment that is unstable under semantically equivalent evaluation protocols is often unreliable, whereas stability alone is not sufficient evidence of correctness.}
Concretely, if an LLM judge produces different rankings when (i)~the rubric is paraphrased, (ii)~the response order is swapped, or (iii)~the judgment is re-sampled at nonzero temperature, the resulting verdict is less trustworthy.
The converse does not hold: a judge can be stably wrong, applying the same erroneous heuristic (e.g., a length or fluency preference) under every protocol variant~\citep{xu2026inside}.
We call this property \emph{protocol stability}, and we show that protocol \emph{instability} is a useful sufficient warning signal in competence regimes where the judge is at least moderately accurate---and we characterize precisely where it fails.

We organize our study around three questions: \emph{(RQ1)} does per-instance protocol stability predict judge error? \emph{(RQ2)} does fusing protocol-stability signals improve the risk--coverage tradeoff over standard confidence and single-signal baselines at a matched budget? \emph{(RQ3)} in which capability and task regimes do these signals help, and where do they break down?

Building on this insight, we introduce \textbf{CARE-Judge} (Calibrated, Abstaining, and Robust Evaluation), which (i)~collects multi-signal protocol-stability features---rubric perturbation, position-swap consistency, and stochastic self-consistency---for each pairwise judgment; (ii)~learns a correctness calibrator $\hat{p}(\text{correct}\mid\text{features})$ on a training split; (iii)~selects an acceptance threshold on a disjoint calibration split by an empirical prefix search certified with an exact Clopper--Pearson bound~\citep{jung2025trust,clopper1934use}; and (iv)~reports selective risk, coverage, and empirical risk-violation rates on a held-out test split.

We validate CARE-Judge across three judge models spanning a wide capability range (Qwen2.5-1.5B, DeepSeek-V4, GPT-5.5) and four benchmarks covering objective correctness (JudgeBench~\citep{tan2024judgebench}), reward-model preference (RewardBench~\citep{lambert2024rewardbench}), summarization preference (TL;DR~\citep{stiennon2020learning}), and open-ended human preference (LMArena~\citep{zheng2023judging}), totaling over 26{,}000 pairwise evaluations.
Our experiments answer the three research questions as follows:

\begin{itemize}
    \item \textbf{(RQ1) Protocol instability predicts judge error.} Rubric perturbation and position-swap consistency produce accuracy gaps of up to 27 percentage points (e.g.\ +27.4pp on DeepSeek/RewardBench) between stable and unstable subsets, consistently across datasets where the judge is at least moderately accurate ($\geq 60\%$). Where the judge is near chance, the same signals lose their predictive sign.
    \item \textbf{(RQ2) Multi-view fusion improves risk--coverage.} At a matched six-call budget, the fused calibrator attains the best or tied-best AUROC in 9 of 12 judge--dataset pairs, beating base confidence, self-consistency, rubric stability, the SCOPE-style bidirectional signal, and a Trust-or-Escalate agreement baseline; e.g., on DeepSeek/TL;DR it lifts pooled accepted accuracy from 67.9\% to 84.1\% under a 15\% risk budget, and adds up to +7.3pp AUROC over the best single signal on DeepSeek-V4.
    \item \textbf{(RQ3) Stability is not correctness.} When Qwen2.5-1.5B (49.4\% on JudgeBench) judges, its verdicts are \emph{stable but wrong}: the signals reverse sign and CARE-Judge abstains on the whole set. We treat this as evidence that protocol stability requires a minimum judge competence, not as a headline success.
\end{itemize}

Our contributions are:
\begin{enumerate}
    \item \textbf{Protocol-stability uncertainty.} We introduce a per-instance uncertainty representation based on invariance under semantically equivalent evaluation protocols---rubric paraphrases, response permutations, and stochastic resampling---and, in particular, are the first to use rubric-paraphrase instability as a per-instance acceptance signal (Section~\ref{sec:method}).
    \item \textbf{Multi-view correctness calibration.} We show that fusing these complementary signals ranks correct versus incorrect verdicts better than standard confidence and single-signal baselines at a matched inference budget, when combined with the established selective-risk apparatus (Section~\ref{sec:experiments}).
    \item \textbf{Risk--coverage characterization across capability regimes.} We map when protocol stability yields useful coverage and when it fails because the underlying judge lacks minimum competence, turning the Qwen--JudgeBench failure into a concrete boundary condition (Section~\ref{sec:experiments}).
\end{enumerate}
We do not claim the risk-control machinery as novel; it is adopted from prior selective-judging work~\citep{jung2025trust,badshah2026scope}. Code and data will be released upon acceptance.
