% sections/related-work.tex

\section{Related Work}
\label{sec:related}

\paragraph{LLM-as-a-Judge.}
The use of LLMs as automated evaluators has grown rapidly.
\citet{zheng2023judging} introduced Chatbot Arena and MT-Bench, establishing pairwise LLM judging as a scalable evaluation paradigm.
\citet{li2024llmsasjudges} provided a comprehensive survey of LLM-based evaluation methods, identifying biases including position bias, length bias, and self-enhancement bias.
\citet{wang2023faireval} proposed mitigating position bias via multi-turn debate and swap-and-compare.
\citet{wei2024systematic} systematically evaluated LLM-as-a-judge prompt sensitivity, finding that rubric template diversity affects alignment scores.
Recent work also cautions that surface agreement need not reflect substantive quality: \citet{xu2026inside} give a mechanistic-interpretability account showing that judge biases arise from shared internal heuristics, which directly motivates our position that stability is a warning signal but not a correctness guarantee.
\citet{li2025preference} further document preference leakage as a contamination mode in LLM-as-a-judge pipelines, underscoring the need for strictly disjoint data splits when calibrating any judge-reliability estimator.

\paragraph{Prompt-Sensitivity, Consistency, and Per-Instance Judge Reliability.}
A growing body of work studies how LLM outputs shift under semantically equivalent prompt, rubric, or ordering perturbations, and how such shifts bear on \emph{per-instance} reliability. \citet{hua2025flaw} ask whether paraphrase sensitivity is a flaw or an artifact; \citet{bryan2025order} isolate the input-order effect; \citet{anghel2025diagnosing} build a pairwise meta-evaluator diagnosing judge instability; and \citet{huang2026prompt} use prompt perturbation to stabilize pairwise evaluation. Closer to the per-instance regime, \citet{gupta2026diagnosing} diagnose judge reliability with conformal prediction sets, \citet{choi2026diagnosing} apply item-response theory, and \citet{li2025autoprompt} ensemble prompt variants to raise accuracy. \citet{ling2025abstention} caution that abstention can itself be a prompt artifact---a caveat we address via competence-gating (Section~\ref{sec:exp:stablewrong}).

We do \emph{not} claim to be the first to observe protocol sensitivity or to use a consistency signal; we differ in \emph{use} and \emph{scope}. In use, prior work diagnoses judges globally (\citealp{hua2025flaw,bryan2025order,anghel2025diagnosing}), builds per-instance reliability \emph{sets} for analysis (\citealp{gupta2026diagnosing,choi2026diagnosing}), or ensembles prompts to \emph{improve accuracy} (\citealp{li2025autoprompt,huang2026prompt}); we instead fuse three perturbation families into a single calibrated accept/abstain decision under an explicit risk budget. In scope, none characterizes the competence-gated asymmetry we report---instability predicts error above a competence threshold and inverts below it.

\paragraph{Selective Prediction and Abstention.}
Selective classification~\citep{geifman2017selective,el2010foundations} enables a predictor to abstain on uncertain inputs.
\citet{wen2025know} surveyed abstention in LLMs, covering verbalized confidence, self-consistency, and retrieval-based approaches.
\citet{krishnan2024enhancing} proposed uncertainty-calibrated fine-tuning for LLM trust, using last-layer hidden states.
Our work differs by focusing on \emph{evaluation-protocol stability} as the uncertainty source, which is specific to the judging task and does not require model internals.

\paragraph{Risk-Controlled Selective Judging.}
Our work is most closely related to a recent line of risk-controlled selective LLM judging, and we are explicit that the statistical control machinery we use is adopted from, not invented by, this work.
\citet{jung2025trust} (Trust-or-Escalate) proposed cascaded selective evaluation for LLM judges with human-agreement guarantees, using Simulated Annotators as the confidence source together with confidence-threshold calibration and Clopper--Pearson risk control.
Concurrently, \citet{badshah2026scope} (SCOPE) proposed selective conformal pairwise judging, aggregating the probabilities of both response orderings into a permutation-invariant score and deriving risk--coverage guarantees on benchmarks including MT-Bench.
\citet{sheng2025analyzing} likewise applied conformal prediction to quantify LLM-as-a-judge uncertainty through interval evaluations.
CARE-Judge uses a comparable selective-risk apparatus---selecting an acceptance threshold on a held-out calibration split via an exact Clopper--Pearson upper bound---rather than proposing new statistical guarantees; we treat this apparatus as an inherited component and are careful (Section~\ref{sec:method}) to describe exactly the empirical, calibration-conditional form of risk control we use.
Our contribution is orthogonal and lies in the \emph{confidence signal}: whereas Trust-or-Escalate relies on simulated-annotator agreement and SCOPE on bidirectional preference probability alone, we combine rubric-paraphrase instability, position-swap inconsistency, and stochastic self-consistency into a single learned calibrator, and we characterize the judge-capability regimes in which such protocol-stability signals help or fail.
Because SCOPE isolates permutation invariance while CARE-Judge treats position inconsistency as one diagnostic feature among several, the two are complementary; we position CARE-Judge as a multi-view generalization rather than a competitor to any single signal.

\paragraph{Judge Benchmarks.}
\citet{tan2024judgebench} introduced JudgeBench, a benchmark of challenging response pairs with objective correctness labels across knowledge, reasoning, math, and coding.
\citet{lambert2024rewardbench} proposed RewardBench for evaluating reward models.
We use JudgeBench and RewardBench as hard, objectively-labeled benchmarks and TL;DR~\citep{stiennon2020learning} and LMArena~\citep{zheng2023judging} as subjective-preference benchmarks, covering both correctness and preference evaluation.

\paragraph{Rubric-Based Evaluation.}
\citet{hong2026from} studied aligning LLM judges with human scoring rubrics.
\citet{anghel2025pearl} proposed a rubric-driven multi-metric evaluation framework.
These works treat rubrics as fixed evaluation criteria to be aligned with; in contrast, we treat \emph{rubric sensitivity}---the degree to which a judge's verdict changes under semantically equivalent rubric paraphrases---as an uncertainty signal.
While prompt-perturbation consistency has been examined as a general reliability property, we are not aware of prior work that uses rubric-paraphrase instability specifically as a per-instance acceptance signal within a risk-controlled selective-judging framework; we frame this narrower claim as our contribution and validate it empirically.

\paragraph{Cost-Aware Cascades.}
Routing easy queries to cheap models and escalating hard ones is a well-established cost-reduction strategy; \citet{zellinger2025costsaving} study LLM cascades with early abstention and formal cost analysis.
Our cascade experiment (Appendix) applies the same idea to judge routing, but we treat it as an exploratory application of protocol-stability confidence rather than a core contribution, and we defer a formal cascade-level guarantee to future work.
